\documentclass[12pt]{book}
\usepackage[ansinew]{inputenc}
%\usepackage{A4}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage[version=3]{mhchem}
\usepackage{siunitx}
\usepackage{hyperref}
\usepackage{onlyamsmath}
%\usepackage{fancyhea,amstex}
%\usestylefile{fancyhea}
%\pagestyle{fancyplain}
%\addtolength{\headwidth}{\marginparsep}
%\addtolength{\headwidth}{\marginparwidth}
\renewcommand{\chaptermark}[1]
           {\markboth{#1}{#1}}
\renewcommand{\sectionmark}[1]
             {\markright{\thesection\ #1}}
%\lhead[\fancyplain{}{\bfseries\thepage}]
%                    {\fancyplain{}{\bfseries\rightmark}}
%\rhead[\fancyplain{}{\bfseries\leftmark}]
%                   {\fancyplain{}{\bfseries\thepage}}
%\cfoot{}





\title{Surface Physics/Chemistry}
\author{Ib Chorkendorff\\ Interdisciplinary Research Center for Catalysis (ICAT)\\ and\\ Center for Atomic-scale Materials Physics (CAMP)\\  at\\ Department of Physics, Building 307\\
Technical University of Denmark\ DK-2800 Lyngby\\ Denmark}
\pagenumbering{roman}
\date{ }

%\includeonly{chap12}
\begin{document}
\maketitle


\addcontentsline{toc}{section}{Contents}
             \tableofcontents

%\section*{Preface}
%This set of notes was made for the course``Surface Physics/Chemistry 10229''. Part I containing Chapter 1-11 covers the fundamental aspects and methods of Surface Science. They are intended to introduce the reader to the most used methods of the field and are considered a prerequsite for understanding the literature.  In the second part the notes are focussed on the fundamental step of gas surface interaction. Here phenomena like gas adsorption, desorption and surface reaction and dynamics are described in detail and it is demonstrated how the priviously described methods can  be used to investigate various surface processes. In the third part are example given how the earlier chapters can be used in relation to heterogeneous catalysis. 
  
\newpage
\pagenumbering{arabic}

%\include{chapA}
\include{chap123}
\include{chap4}
\include{chap5}
\include{chap6}
\include{chap7} 
\include{chap8}
\include{chap9}
\include{chap10}
\include{chap11}

%\include{chapB}

%\include{chap12}

%\include{chap13}
%\include{chap14}
%include{chap15}
%\include{chap16}
%\include{chap17}



\begin{thebibliography}{99}
\addcontentsline{toc}{chapter}{Bibliography}

\bibitem{Zangwill} A. Zangwill, ``Physics at Surfaces'', Cambridge University Press, Cambridge (1988) ISBN 0-521-34752-1.

\bibitem{Somorjai} G. Somorjai,``Introduction to Surface Chemistry and Catalysis'',Wiley and Sons, New York (1994) ISBN 0-471-03192-5.

\bibitem{Hudson} J. B. Hudson ``Surface Science'', Wiley and Sons, New York (1998), ISBN 0-472-25239-5.

\bibitem{Niemantsverdriet} J. W. Niemantsverdriet, ``Spectroscopy in Catalysis'', VCH, Weinheim (1995), ISBN 3-527-28726-4.

\bibitem{Ertl} G. Ertl and J. Kuppers, ``Low energy electrons and Surface Chemistry'' VCH Weinheim (1985), ISBN 3-527-26056-0.

          \bibitem{briggs} D. Briggs and M. P. Seah, ``Practical
          Surface Analysis'', Wiley (1990) ISBN 0 471 92081 9.

             \bibitem{feldman}L. Feldman and J. W.
          Mayer,``Fundamentals of Surface and Thin Film Analysis'',
          North-Holland (1986) ISBN 0 444 00989 2.


             \bibitem{woodruff} D. P. Woodruff and T. A.
          Delchar,``Modern techniques of surface science'', Cambridge
          University Press (1986) ISBN 0 521 35719 5.

%Chapter 2 ***********************************************



          \bibitem{roth} A. Roth, ``Vacuum Technology'', North Holland
          (1990), ISBN 0 444 86027 4.

	\bibitem{Somorjai_two_dimensions} G. Somorjai,``Chemistry in two dimensions'', Cornell University Press, (1981).

             \bibitem{Feynman} R. P. Feynman, R. B. Leighton, and M.
          Sands,``The Feynman Lectures'' Vol II. Addison-Wesley
          Publishing Company (1975).



             \bibitem{siegbahn1}  K.  Siegbahn,  C.  N.  Nordling,   K.
          Hamrin, J. Hedman, G. Johansson, T. Bergmark, S. E. Karlson,
          I. Lindgren, and B. Lindberg. ``ESCA, Atomic, Molecular, and
          Solid  State  Structure  Studied  by   Means   of   Electron
          Spectroscopy'', Almqvist and Wiksells, Uppsala, Sweeden, (1967).



             \bibitem{krause} Krause and Ferreria, (1975).

          \bibitem{chorkendorff}  I.  Chorkendorff,  J.  Onsgaard,   J.
          Schmidt-May, and R. Nyholm, Surf. Sci. 160 (1985) 587.


          \bibitem{gerken} F. Gerken, PhD Thesis,  HASYLAB,  Hamburg,
          Germany, (1982).

          \bibitem{shirley} D. A. Shirley, Topics in Applied  Physics,
          27, (19xx).



             \bibitem{wertheim} G. K. Wertheim and L.  R.  Walker,  J.
          Phys. F: Metal Phys, 6 (1976) 2297.



          \bibitem{doniach} S. Doniach and  M.  Sunjic,  J.  Phys.  C:
          Solid St. Phys, 3 (1970) 285.

          \bibitem{perkin} C. D. Wagner, W. M. Riggs, L. E. Davis,  J.
          F. Moulder,  and  G.  E.  Muilenberg,  ``Handbook  of  X-ray
          Photoelectron Spectroscopy'', Perkin-Elmer Corporation.

          \bibitem{ghosh} P. K. Ghosh, ``Introduction to Photoelectron
          and Auger Spectroscopy'', Wiley-Interscience Publishers, New
          York, (1983).

          \bibitem{shofield} J. H. Schofield, J. Electron Spectrosc. 8
          (1976) 129.


             \bibitem{chorkendorff2} I. Chorkendorff, I. Alstrup,  and
          S. Ullmann., Surf. Sci. 227 (1990) 291.

% Chapter 5 UPS*********************************************


\bibitem{Plummer} E. W. Plummer and W. Eberhart, Adv. Chem. Phys. 49 (1982) 533.
\bibitem{Andersen} J. N. Andersen I. Chorkendorff, J. Onsgaard, J. Ghiejsen, R. L. Johnson, and F. Grey, Phys. Rev. B 37 (1988) 4809.
\bibitem{Neddermeyer} P. Heimann, J. Hermanson, J. Miosga, and H. Neddermeyer Phys. Rev. B 20 (1979) 3059.
\bibitem{Kevan} S. D. Kevan, Phys. Rev. Lett. 50 (1983) 526.
\bibitem{Berglund} C. N. Berglund and W. E. Spicer, Phys. Rev. 136 A (1964) 1030 ibid. 1044.
\bibitem{Gustafson} T. Gustafson and E. W. Plummer, in Photoemission and Electronic Properties of Surfaces; Ed B. Feuerbacher, B. Fitton, and R. F. Willis,Wiley, Chicchester, p53 (1978).

%chapter 6 AES**********************************************


         \bibitem{wentzel} G. Wentzel, Z. Phys. 43 (1927) 524.

             \bibitem{handbook}  L.  E.  Davis,   N.   C.   Macdonald,
          P. W. Palmberg, G. E. Riach. and R. E. Weber, ``Handbook  of
          Auger Spectroscopy'', Perkin-Elmer Corporation.

             \bibitem{chorkendorff3} I.  Chorkendorff,  J.  N.  Russel
          Jr., and J. T. Yates Jr., J. Chem. Phys 86 (1987) 4692.

             \bibitem{briggs2} D. Briggs and M. P. Seah, ``Practical
          Surface Analysis'' Volume II, Wiley (1990) ISBN.




          \bibitem{sigmund}  P.  Sigmund,   ``Sputtering   by   Ion
          Bombardment;  Theoretical  Concepts''   in   Sputtering   by
          Particle Bombardment I, Editor R. Behrisch.  Springer-Verlag
          New York, (1982). Chapter 2.

          \bibitem{sigmund2} P. Sigmund, Phys. Rev. 184 (1969) 383.

          \bibitem{behrisch}             R. Behrisch.  Sputtering   by
          Particle Bombardment I and  II,  Springer-Verlag  New  York,
          (1982). Chapter 2.

          \bibitem{zalar} A. Zalar, Thin Solid Films 124, (1985) 223.

          \bibitem{hoffman} S. Hoffmann, Chapter 4 in D. Briggs and M.
          P.  Seah,  ``Practical  Surface  Analysis''Volumen  I,  Wiley
          (1990) ISBN 0 471 92081 9.

          \bibitem{alstrup} N. C. Alstrup, PhD Thesis, DTU (1993).
% Chapter 8 ****************************************************'
\bibitem{Ibach}  H. Ibach and D. L. Mills, ``Electron Energy Loss Spectroscopy and Surface Vibrations'' Academic Press, New York (1982).
\bibitem{Bruckmann} T. Bruckmann, S. Lehrwald, G. Kisters, and H. Ibach, (1990).
\bibitem{Formate} P. A. Taylor, P. Rasmussen, C. V. Ovesen, P. Stoltze, and I. Chorkendorff, Surf. Sci. 261 (1992) 191.
\bibitem{Bradshaw} A. M. Bradshaw, Surf. Sci. 11/12 (1982) 712.
\bibitem{Barnes} M. R. Barnes and R. F. Willies, Phys. Rev. Lett. 41 (1978) 1729.
\bibitem{Bailey}  S. Bailey, G. Forment, J. W. Snoeck, and K. C. Waugh, Catal. Lett. 30 (1995) 99.


% Chapter 9 LEED***********************************************

\bibitem{Kittel} C. Kittel, Elementary Solid State Physics: A Short Course, Wiley, New York, 1962.
\bibitem{Ashcroft} N. W. Ashcroft and N. D. Mernib, Solid State Physics, Holt-Saunders, Philadelphia, 1976.
\bibitem{Davidson}  C. J. Davidson and L. H. Germer, Phys. Rev. 30 (1927) 705.
\bibitem{Pendry} J. B. Pendry, Low Energy Electron Diffraction, Academic Press, New York 1974.
\bibitem{Hove} M. A. van Hove and  S. Y. Tong, Surface Crystallography by LEED, Springer, Berlin 1979.
\bibitem{Heinz} K. Heinz and K. Muller, in Structual Studies of Surfaces, Springer Tracts in Modern Physics, vol 91. Springer Berlin 1982.
\bibitem{HonCu} P. A. Taylor, P.B. Rasmussen, C. V. Ovesen, P. Stoltze, and I. Chorkendorff, Surf. Sci. 261 (1992) 191.
\bibitem{Shiff} L. I. Schiff, ``Quantum Mechanics'' McGraw-Hill Tokyo (1968).
\bibitem{Merzbacher} E. Merzbacher, ``Quantum Mechanics'' Wiley New York (1970).




% Chapter 10***********************************
\bibitem{Roher1} G. Binnig, H. Roeher, C. Gerber, and E. Weibel, Appl. Phys. Lett. 40 (1982) 178.
\bibitem{Roher2} G. Binnig, H. Roeher, C. Gerber, and E. Weibel, Phys. Rev. Lett. 49 (1982) 57.
\bibitem{Roher3} G. Binnig, H. Roeher, C. Gerber, and E. Weibel, Phys. Rev. Lett. 50 (1983) 120.
\bibitem{STM} Scanning Tunneling Microsopy I, II, III, Springer Series in Surface Science Ed: R. Wiesendanger and H. J. Guntherrodt,  (1993)  Berlin, ISBN 3-540-56317-2.
\bibitem{Tersoff} J. Tersoff and D. R. Hamann, Phys. Rev. Lett 50 (1983) 1998.
\bibitem{Bardeen} J. Bardeen, Phys. Rev. Lett. 6, (1961) 57.
\bibitem{Lang} N. Lang, Chapter 2 in Scanning Tunneling Microsopy  III, Springer Series in Surface Science, Ed: R. Wiesendanger and H. J. Guntherrodt,  (1993)  Berlin, ISBN 3-540-56317-2.
\bibitem{Luigi} M. L Colaianni and I. Chorkendorff, Phys. Rev. B 50 (1994) 8798.
\bibitem{Besenbacher1} F. Besenbacher, Rep. Prog. Phys. 59 (1996) 1737.
\bibitem{Besenbacher2} F. Besenbacher, Chapter 10 in``Techenical Physics of Solid Surfaces and Heterogeneous Catalysis", Ed. D. King and  D. P. Woodruff.
\bibitem{Besenbacher3} L. Pleth Nielsen, PhD Thesis from University of \AA rhus, Denmark (1996).
\bibitem{Science}  F. Besenbacher, I. Chorkendorff, B. S. Clausen, B. Hammer, A. M. Molenbroek, and J. K. N\o rskov, Science 279 (1998) 1913.
\bibitem{Morten} M. \O. Pedersen, I. A. Bonicke, E. L\ae gsgaard, I. Stensgaard, A. Ruban, J. K. N\o rskov, and F. Besenbacher, Surf. Sci. 387 (1997) 86.
\bibitem{Datye} A. K. Datye and  N.J. Long, Ultramicroscopy 25 (1988) 203.
\bibitem{carbon} J. R. Rostrup-Nielsen, {\em Catalytic Steam Reforming} Catalysis- Science and Technology. Edited by J. R. Anderson and M. Boudart, Springer-Verlag, Berlin 1984, ISBN 3-540-12665-1.

%Chapter 11************************************
\bibitem{Yates} S. M. Gates, J. N. Russel Jr., and  J. T. Yates Jr., Surf. Sci. 159 (1985) 233.
\bibitem{Yates1} H. R. Siddiqui, X. Guo, I. Chorkendorff, and J. T. Yates Jr., Surf. Sci. 191 (1987) L813.
\bibitem{Rasmussen} P. B. Rasmussen and I. Chorkendorff, Surf. Sci. 248 (1991) 35.
\bibitem{Yates2} J. N. Russel Jr., I. Chorkendorff, and J. T. Yates Jr., Surf. Sci. 183 (1987) 316.


%Chapter 11************************************
\bibitem{Brongersma} ???
\bibitem{Norskovlang} N. D. Lang and J. K. N\o rskov, Physica Scripta 6 (19983) 15.
\end{thebibliography}
\end{document}

